---
title: "Reflections on Data Science"
format: html
---

# What is being a data scientist about?

Being a data scientist is about **turning questions into evidence**.  
It involves clarifying problems, preparing and analyzing data, creating and applying models,  
and communicating insights so that others can act on them.  
It’s detective work, engineering, and storytelling combined.

Some common activities include:
- Wrangling and data cleaning and transformation
- Building and validating models
- Sharing results in clear, reproducible reports
- Thinking about ethics and the impact of analysis

# Data scientists vs. statisticians

**Similarities**
- Both rely on data, probability, and models.
- Both value reproducibility and critical thinking.

**Differences**
- Statisticians often emphasize *theory, inference, and study design* and usually focus on the event modeling and the relationship between the event/target and input (independent) variables, and the sampling process as well.  
- Data scientists often emphasize *end‑to‑end workflows, coding, and deployment*, focusing on agile models that fit product needs and challenging production models with **challenger** approaches.  

**My view**  
I see myself as combining both: I like the theoretical rigor of statistics  
and the practical, tool‑driven mindset of data science.

# Exploring a density object in R

We will now create a **kernel density estimate** of `iris$Petal.Width`.

First, we create the object:

```{r}
# Step 1: Create the object
y <- density(iris$Petal.Width)
```

Next, we look inside the object:

```{r}
# Step 2: Look inside the object
class(y)
typeof(y)
str(y)
```

Finally, we plot the density.  
The code will be hidden in the HTML:

```{r}
#| echo: false
# Step 3: Make the plot
plot(
  y,
  main = "Kernel Density of iris$Petal.Width",
  xlab = "Petal Width (cm)",
  ylab = "Estimated Density"
)
```